Classification and Regression Trees and Beyond

### 2020-06-30 pkg/partykit/inst/ULGcourse-2020/slides_tree.Rnw
changed title changed author order streamlined LaTeX header
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@3019 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: b392afeb40779fe25199dc93b3e0adc981db2198

\documentclass[11pt,t,usepdftitle=false,aspectratio=169]{beamer}
\usetheme[nototalframenumber,url,license]{uibk}

\title{Tree models}
\subtitle{Supervised Learning: Algorithmic Modeling}
\author{Achim Zeileis, Lisa Schlosser}
\setbeamerfont{url}{size*={11.5pt}{13pt},series=\mdseries}
%\URL{http://www.partykit.org/partykit/}
\renewcommand{\headerimage}[1]{%
   \IfStrEqCase{#1}{%
      {1}{%
         \gdef\myheaderimageid{#1}%
         \gdef\myheaderimageposition{nw}%
         \gdef\myheaderimage{forest.jpg}%
      }}[%
         \gdef\myheaderimageid{1}%
         \gdef\myheaderimageposition{nw}%
         \gdef\myheaderimage{forest.jpg}%
      ]%
}
\headerimage{1}

%\graphicspath{{Figures/}}
\usepackage[utf8]{inputenc}

\setbeamertemplate{caption}{\insertcaption} 
%% includes a replacement for \usepackage{Sweave}
% \usepackage{Sweave}
\usepackage{changepage}
\usepackage{amsmath,tikz}
\usepackage{calc}
\usepackage{graphicx}
\usetikzlibrary{positioning,shapes,arrows,decorations.pathreplacing,calc,automata,mindmap,trees,tikzmark,decorations.pathreplacing}
\usepackage{color}
\usepackage{xcolor}
\usepackage{fancyvrb}
%\usepackage[cal=boondoxo]{mathalfa}
%\graphicspath{{plots/}}
\newcommand{\argmax}{\operatorname{argmax}\displaylimits}

\setbeamertemplate{subsection page}
{\begin{center}
\vspace{2cm}
% {\Large \insertsection\par}
% \vspace{0.5cm}
Section~\thesubsection\par
\vspace{0.5cm}
{\LARGE \insertsubsection\par}
\end{center}}
\AtBeginSubsection{\frame{\subsectionpage}}

%% colors
% \definecolor{HighlightOrange}{rgb}{0.9490196,0.5725490,0.0000000}
% \definecolor{HighlightBlue}{rgb}{0.4784314,0.7490196,0.9803922}
% \definecolor{forestred}{RGB}{206,73,81}
% \definecolor{treegreen}{RGB}{0,143,0}
% \definecolor{lightblue}{RGB}{34,151,230}
% \definecolor{lightorange}{RGB}{255,165,0}

%\SweaveOpts{engine=R, eps=FALSE, keep.source=TRUE}
\SweaveOpts{engine=R, eps=FALSE, echo=FALSE, results=hide, keep.source=TRUE}

@@

set.seed(7)

invisible(.Call(grDevices:::C_palette, grDevices::hcl(
  h = c(0,   5, 125, 245, 195, 315,  65,   0),
  c = c(0, 100,  90,  85,  63, 105,  90,   0),
  l = c(0,  55,  75,  60,  82,  48,  80,  65)
)))

# setwd("~/Documents/ULG/slides_tree/")
library("disttree")
library("Formula")
library("latex2exp")
library("AER")
#library("gamlss")
#library("lattice")
#library("crch")
#library("RainTyrol")
#library("parallel")
library("rpart")
#library("gamlss.cens")
#gen.cens(NO, type = "left")
#library("ggplot2")
#theme_set(theme_bw(base_size = 18))
library("MASS")
#library("Ecdat")
#library("wooldridge")
library("partykit")
#library("RWeka")
#library("evtree")
library("coin")

## HCL palette
# pal <- hcl(c(10, 128, 260, 290, 50), 100, 50)
# names(pal) <- c("forest", "tree", "gamlss", "gamboostLSS", "EMOS")
# 
# pallight <- hcl(c(10, 128, 260, 290, 70), 100, 50, alpha = 0.25)
# names(pallight) <- c("forest", "tree", "gamlss", "gamboostLSS", "EMOS")
# 
# transpgray <- rgb(0.190,0.190,0.190, alpha = 0.2)
@

<<data>>=
data("Titanic", package = "datasets")
ttnc <- as.data.frame(Titanic)
ttnc <- ttnc[rep(1:nrow(ttnc), ttnc$Freq), 1:4]
names(ttnc)[2] <- "Gender"
# ttnc <- transform(ttnc, Treatment = factor(Gender == "Female" | Age == "Child", 
#                                            levels = c(FALSE, TRUE),
#                                            labels = c("Male&Adult", "Female|Child")))
##   labels = c("Normal\n(Male&Adult)", "Preferential\n(Female|Child)")
@

@@

\section{Tree models}

@@

\end{frame}
\begin{frame}[fragile]
\frametitle{Motivation}
{\small Wages:  \quad $log(wage) \; \sim  \; education + experience + age + ethnicity + gender + union$}
\begin{center}
\setkeys{Gin}{width=0.99\textwidth}
<<ctree_Wages1985, echo=FALSE, fig=TRUE, width=13>>=
data("CPS1985")
plot(ctree(log(wage)~education+experience+age+ethnicity+gender+union,data=CPS1985,
           control = ctree_control(alpha = 0.01)))
@
\end{center}
\end{frame}

\begin{frame}[fragile]
\frametitle{Motivation}
{\small Wages:  \quad $log(wage) \; \sim  \; education \ | \ experience + age + ethnicity + gender + union$}
\begin{center}
\setkeys{Gin}{width=0.9\textwidth}
<<lmtree_Wages1985, echo=FALSE, fig=TRUE, width=11>>=
data("CPS1985")
plot(lmtree(log(wage)~education|experience+age+ethnicity+gender+union,data=CPS1985))
@
\end{center}
\end{frame}

@@

% \begin{frame}[fragile]
% \frametitle{Motivation}
% {\small Wage data:  \quad $log(wage) \; \sim  \; education + experience + ethnicity + region + parttime$}
% \begin{center}
% \setkeys{Gin}{width=0.99\textwidth}
% <<ctree_Wages1988, echo=FALSE, fig=TRUE, width=13>>=
% data("CPS1988")
% ## ctree
% plot(ctree(log(wage)~education+experience+ethnicity+region+parttime,data=CPS1988,
%            control = ctree_control(alpha = 0.01, minsplit = 7000)))
% @
% \end{center}
% \end{frame}

>>>>

\documentclass[11pt,t,usepdftitle=false,aspectratio=169]{beamer}
\usetheme[nototalframenumber,license]{uibk}

\title{Classification and Regression Trees and Beyond}
\subtitle{Supervised Learning: Algorithmic Modeling}
\author{Lisa Schlosser, Achim Zeileis}

%% forest header image
\renewcommand{\headerimage}[1]{%
   \IfStrEqCase{#1}{%
      {1}{%
         \gdef\myheaderimageid{#1}%
         \gdef\myheaderimageposition{nw}%
         \gdef\myheaderimage{forest.jpg}%
      }}[%
         \gdef\myheaderimageid{1}%
         \gdef\myheaderimageposition{nw}%
         \gdef\myheaderimage{forest.jpg}%
      ]%
}
\headerimage{1}

%% custom subsection slides
\setbeamercolor*{subsectionfade}{use={normal text},parent={normal text},fg=structure.fg!30!normal text.bg}
\AtBeginSubsection[]{%
  \begin{frame}[c]
    \begin{center}
      \usebeamercolor[fg]{subsectionfade}
      \Large \insertsection \\[2ex]
      \usebeamercolor[fg]{structure}
      \huge\bfseries\insertsubsection
    \end{center}
  \end{frame}
}

%% for \usepackage{Sweave}
\SweaveOpts{engine=R, eps=FALSE, echo=FALSE, results=hide, keep.source=TRUE}

@@

library("rpart")
library("partykit")
library("coin")

set.seed(7)
@

<<Titanic-data>>=
data("Titanic", package = "datasets")
ttnc <- as.data.frame(Titanic)
ttnc <- ttnc[rep(1:nrow(ttnc), ttnc$Freq), 1:4]
names(ttnc)[2] <- "Gender"
@

@@

\section{Classification and regression trees and beyond}

@@

\end{frame}
\begin{frame}[fragile]
\frametitle{Motivation}
{\small Wages:  \quad $log(wage) \; \sim  \; education + experience + age + ethnicity + gender + union$}
\begin{center}
\setkeys{Gin}{width=0.99\textwidth}
<<ctree_Wages1985, echo=FALSE, fig=TRUE, width=13>>=
data("CPS1985", package = "AER")
plot(ctree(log(wage)~education+experience+age+ethnicity+gender+union,data=CPS1985,
           control = ctree_control(alpha = 0.01)))
@
\end{center}
\end{frame}

\begin{frame}[fragile]
\frametitle{Motivation}
{\small Wages:  \quad $log(wage) \; \sim  \; education \ | \ experience + age + ethnicity + gender + union$}
\begin{center}
\setkeys{Gin}{width=0.9\textwidth}
<<lmtree_Wages1985, echo=FALSE, fig=TRUE, width=11>>=
data("CPS1985", package = "AER")
plot(lmtree(log(wage) ~ education | experience + age + ethnicity + gender + union, data = CPS1985))
@
\end{center}
\end{frame}

@@

% \begin{frame}[fragile]
% \frametitle{Motivation}
% {\small Wage data:  \quad $log(wage) \; \sim  \; education + experience + ethnicity + region + parttime$}
% \begin{center}
% \setkeys{Gin}{width=0.99\textwidth}
% <<ctree_Wages1988, echo=FALSE, fig=TRUE, width=13>>=
% data("CPS1988", package = "AER")
% ## ctree
% plot(ctree(log(wage)~education+experience+ethnicity+region+parttime,data=CPS1988,
%            control = ctree_control(alpha = 0.01, minsplit = 7000)))
% @
% \end{center}
% \end{frame}
